---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

## forestdata

`forestdata` is an R package for downloading forestry and land use data. This package aims to provide users with easy access to various datasets related to forestry and land use.

## Installation

You can install the development version of `forestdata` from GitHub using the `pak` package:

``` {r, eval=FALSE}
pak::pak("Cidree/forestdata")
```

## Functions

In the following table, you can find a summary of the available functions, a short description, and the object type of the output.

```{r, echo=FALSE}
# Load necessary libraries
library(tibble)
library(gt)

# Create a tibble with the specified columns
forestdata_info <- tibble(
  Function = c(
    "fd_landcover_copernicus", 
    "fd_landcover_esri",
    "fd_canopy_height",
    "fd_forest_chorological",
    "fd_forest_eutrees4f",
    "fd_forest_extent_glad",
    "fd_forest_spain_mfe50",
    "fd_forest_france",
    "fd_inventory_spain",
    "fd_pathogens_defid2"
  ),
  Description = c(
    "Download data from the Global Land Cover from the Copernicus Global Land Service.",
    "Download data from the ESRI Land Cover Explorer.",
    "Download data from the ETH Global Sentinel-2 canopy height from 2020.",
    "Download data from the European Chorological Maps.",
    "Download data from the EU-Trees4F Database.",
    "Download data for the forest extension from the Global Land Analysis & Discovery.",
    "Download the Spanish Forest Map (MFE50) for a province.",
    "Download the BD Forêt for a French Department (French Forest Map).",
    "Download data for the IFN2, IFN3 or IFN4 for a Spanish province.",
    "Download the Database of European Forest Insect and Disease Disturbances (DEFID)."
  ),
  Output = c(
    "SpatRaster", 
    "SpatRaster", 
    "SpatRaster", 
    "sf", 
    "SpatRaster", 
    "SpatRaster", 
    "sf", 
    "sf", 
    "list", 
    "sf"
  ),
  group = c(
    rep("Land Cover", 2),
    rep("Forest", 7),
    rep("Forest Pathogens", 1)
  )
)

# Render the tibble as a nice table using gt
gt(forestdata_info, groupname_col = "group") |>
  opt_stylize(style = 3, color = "cyan") |> 
  tab_header(
    title    = "{forestdata} Package Functions",
    subtitle = "Download and Process Forestry and Land Use Data"
  ) |>
  cols_label(
    Function = "Function",
    Description = "Description",
    Output = "Output"
  ) |>
  fmt_markdown(
    columns = c(Description)
  ) |>
  cols_align(
    align = "left",
    columns = c(Function, Description, Output)
  ) |>
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12
  ) |> 
  tab_options(
    table.font.name           = "Merriweather",
    column_labels.font.size   = 14,
    column_labels.font.weight = "bold",
    row_group.font.size       = 14,
    row_group.padding         = 10,
    row_group.font.weight     = "bold",
    data_row.padding          = 10
  )
```



## Note

Please read the function documentation carefully. Some datasets may require proper citation when used.
