---
title: "Canopy Height Models"
editor_options: 
  chunk_output_type: console
---

## Introduction

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This page shows how to download Canopy Height Models using the `forestdata` R package. First, we will load some packages to get the data and create a map:

```{r setup, output = FALSE}
library(forestdata)
library(giscoR)
library(ggplot2)
library(sf)
library(terra)
library(tidyterra)
```

The function that we will use to retrieve the data is `fd_canopy_height()`, and it has several useful arguments. First, we can either specify a polygon (`sf` or `SpatVector` object) in the `x` argument, or specify a longitude and latitude (`lon` and `lat`) values for retrieving a tile.

Next, we have the argument `model`, which can be set to:

* `model = "eth"`: to use the ETH Global Sentinel-2 10m canopy height from 2020 ([more information](https://www.research-collection.ethz.ch/handle/20.500.11850/609802)).

* `model = "meta"`: to use the very high resolution canopy height maps (1m) of Meta ([more information](https://www.sciencedirect.com/science/article/pii/S003442572300439X)).

* `model = "amazon"`: to use the very high resolution canopy height maps (5m) of the Amazon Forest ([more information](https://arxiv.org/abs/2501.10600)).

The argument `layer` only applies to the `eth` model. We can retrieve the canopy height model (`layer = "chm"`), the standard deviation (`layer = "std"`) or both layers (`layer = "all"`).

## Meta models

The Global Canopy Height Maps dataset provides a detailed view of tree canopy heights worldwide from 2009 to 2020, focusing on data from 2018 to 2020, which accounts for 80% of the information. Developed by Meta and the World Resources Institute, it serves as a critical reference for monitoring forest structure and assessing carbon stocks, supporting carbon credit schemes. Utilizing advanced satellite imagery and AI models like DiNOv2, it achieves a high precision with a mean absolute error of 2.8 meters, enabling accurate detection of changes in canopy height over time and improving carbon monitoring efforts.

We will create a map of the canopy height using one tile from La Palma Island (Canary Islands, Spain). To do so, we will download the Island boundaries using the `{giscoR}` package:

```{r}
## Download La Palma
palma_sf <- gisco_get_nuts(
  country    = "Spain",
  nuts_level = 3,
  resolution = "03"
) |> 
  filter(NAME_LATN == "La Palma")

## Print
print(palma_sf)
```

Once we have downloaded the study area, we can proceed to download the canopy height model of Meta. We can use the `crop` and `mask` arguments to extract only the area of the island:

```{r}
## Download the data
palma_meta_sr <- fd_canopy_height(
  x     = palma_sf,
  model = "meta",
  crop  = TRUE,
  mask  = TRUE
)

## Print raster
print(palma_meta_sr)
```

The tiles are stored in a [AWS Bucket](https://aws.amazon.com/s3/), which is internally accessed using the `aws.s3` R package. Therefore, you will probably need to install this package if you want to use this function, since it's not imported by `forestdata`:

```{r, eval=FALSE}
pak::pak("aws.s3")
```


We can now explore the height's distribution in La Palma:

```{r}
hist(palma_meta_sr, main = "Distribution of heights", xlab = "Height (m)")
```

Once we have the data, we can generate one map of the canopy height model leveraging the `tidyterra` package:

```{r, fig.width=6, fig.height=8}
ggplot() +
  geom_spatraster(data = palma_meta_sr) +
  scale_fill_whitebox_c(
    name      = "Height (m)",
    palette   = "gn_yl",
    direction = -1
  ) +
  guides(
    fill = guide_legend(
      position       = "top",
      direction      = "horizontal",
      title.position = "top",
      title.hjust    = .5,
      label.position = "bottom",
      label.hjust    = .5
    )
  ) +
  labs(
    title   = "Canopy Height Model in La Palma (Spain)",
    caption = "@ Adrián Cidre | Data Source: Meta CHM | {forestdata} R package"
  ) +
  theme_void(
    base_size = 10
  ) +
  theme(
    panel.background  = element_rect(fill = "#8DA7BE", colour = NA),
    plot.background   = element_rect(fill = "#8DA7BE", colour = NA),
    plot.title        = element_text(
      face   = "bold", 
      hjust  = .5, 
      margin = margin(t = 10, b = 10)
    ),
    plot.caption      = element_text(hjust = .5, margin = margin(b = 10)), 
    legend.key.height = unit(2, "mm"),
    legend.key.width  = unit(7, "mm")
  )
```


















